MLH1 Results Outline

### 2020-02-17 doc/Results.rmd
trying to finalize the Q1.1CO pos section

```{r FEMALE.TABLE, echo=FALSE}

@@

```{r MALE.TABLE, echo=FALSE}

@@

```{r MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
#DF.HetC.MixedModel.HQ

@@

@@

```{r tot.sc.viz4outliers, echo=FALSE}

@@

From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.

@@

skel -- takes the skeletonized midline of the bin objects

# Main Figure

@@

The strain means for females were pretty similar. For female Domesticus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Dom")  )$mean_co` for _M. m. domesticus^WSB^_, _M. m. domesticus^G^_, and _M. m. domesticus$^LEW^_ respectively. In the three Musculus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Musc")  )$mean_co` for _M. m. musculus^ PWD}^_, _M. m. musculus^ SKIVE^_, and _M. m. musculus^KAZ^_ respectively. In the molossinus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Mol")  )$mean_co` for _M. m. molossinus^ MSM^_ and _M. m. molossinus^MOLF^_ respectively.

While in two of the subspecies, the means for males show a greater range across means, notable in the mmmolossinus strains `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Mol")  )$mean_co` for _M. m. molossinus^MSM^_ and  _M. m. molossinus^MOLF^_ respectively. In the musculus the strain means are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Musc")  )$mean_co` for _M. m. musculus^PWD^_, _M. m. musculus^SKIVE^_, _M. m. musculus^KAZ^_, _M. m. musculus^TOM^_, _M. m. musculus^AST^_, and _M. m. musculus^CZECH^_ respectively. In the male Dom strains the means for MLH1 count per cell are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Dom")  )$mean_co`  for  _M. m. domesticus^WSB^_, _M. m. domesticus^G^_, _M. m. domesticus^LEW^_ and _M. m. domesticus^PERC^_ respectively. These

<!-- mouse level means-->
The ranges of mean MLH1 counts per cell are

@@

# Analysis for Evolutionary Patterns

@@

Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
-->s

@@

```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}

@@

#female strain table
#female_mouse.avs_for.analysis
F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,
                        Nmice =length(unique(mouse)),
                        Ncells  = sum(Ncells),
                        mean_co =as.numeric(format(round(mean(mean_co), 5 ), nsmall=3) ),

@@

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se)
    )
F.strain.table
```

@@

In the dom  subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in _M. m. domesticus$^WSB^_
`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` in _M. m. domesticus$^G^_

@@

```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}
#try customize the plot size

@@

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se)
    )
M.strain.table
```

@@

(discription of the nice boxplot figure)

@@

- the ratios of DMC1 : MLH1 counts!!

```{r plot.stage.ttest, echo=FALSE, warning=FALSE}

@@

```{r DMC1_MLH1.cor, echo=FALSE}
#make category averages from AP_mouse_table.HQ
mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
                 .(category),
                 summarize, 
                 mean.MLH1 = mean(mean_co),
                 var.MLH1 = var(mean_co)
)

@@

PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1
```

The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`.

The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.

Why calculate ratios using the mean Leptotene cells -- they represent the a metric for the CO : NCO decision ( DSB / COs = NCO)

ratios are .. not that different (slightly suprisingly)

`r c(WSB_DMC1.Z_CO_ratio, G_DMC1.Z_CO_ratio, MSM_DMC1.Z_CO_ratio, PWD_DMC1.Z_CO_ratio )` for  _M. m. domesticus^WSB^_ , _M. m. domesticus^G^_, , _M. m. molossinus^MSM^_ and  _M. m. musculus^PWD^_ respecitively.

@@

(maybe ignore KAW.L? 1 cell)

@@

Since there is evidence for non-equal variance across strains for zygotene cells, so don't rely on lm()s that estimate the effect of strains.

The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.HighLow.L$p.value`) while the zygotene cells, there is not a significant difference across the high and low groups (value = `r ttest.HighLow.Z$p.value`).

@@

ToDo: chisq tests for difference in proportions, 
- within strain across sex,
- males, across high and low groups
- look at mouse level proportions
-->

<!-- merging the 2 data sets-->

@@

```{r Q1.biv.data_table.4MM.HetC, echo=FALSE, eval=FALSE}

#this block should be in the data set up

#the mouse averages will be used

#assign high and low
#divide musc males into groups
mouse.avs_4MM$Rec.group.alt <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))

mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))

In the chunk above the mouse averages table is made -- may need to add all the extra metrics (IFD, .

@@

Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
summary(Q1.glm_totSC_M2)
#subsp effect, many strain, #male

@@

Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
summary(Q1.glm_totSC_M3)
#LOT OF SIGNIFICant strain effects in this model

@@

```{r Q1.SC.Length_M2, echo=FALSE}

Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)

@@

```{r JUNK.Q1.SC_short_t.test, echo=FALSE, include=FALSE}

#t.tests for a
Q1_short.sc_t.test.WSB <- t.test((Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "WSB") )$mean_SC,
                           (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == "WSB") )$mean_SC)
#sig

#Using the full data set has strong signle but -- the interpertation is complicated
#FULL SC t.test
#run a t.test
wsb.ttest  <- t.test(WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "female"], WSB.Bivalents.DF$chromosomeLength[WSB.Bivalents.DF$sex == "male"])

#run a t.test seperate out by the chrm classes
LEW.ttest  <- t.test(LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "female"], LEW.Bivalents.DF$chromosomeLength[LEW.Bivalents.DF$sex == "male"])
#run a t.test seperate out by the chrm classes
G.ttest  <- t.test(G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "female"], G.Bivalents.DF$chromosomeLength[G.Bivalents.DF$sex == "male"])
#run a t.test seperate out by the chrm classes
PWD.ttest  <- t.test(PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "female"], PWD.Bivalents.DF$chromosomeLength[PWD.Bivalents.DF$sex == "male"])

#run a t.test seperate out by the chrm classes
KAZ.ttest  <- t.test(KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "female"], KAZ.Bivalents.DF$chromosomeLength[KAZ.Bivalents.DF$sex == "male"])
#run a t.test seperate out by the chrm classes
SKIVE.ttest  <- t.test(SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "female"], SKIVE.Bivalents.DF$chromosomeLength[SKIVE.Bivalents.DF$sex == "male"])

@@

To quantify one of the major -- patterns that is different between male and female recombiantion landscapes

In order to test for evolution -- and examine sex differences in the recombination landscape (crossover position bias)

we focus on the foci positions from one crossover (1CO) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference.

```{r Q1show.F1pos_Full, echo=FALSE, eval=FALSE}

@@

As expected -- confirming previous results, the normalized position of single crossovers varies between males and females. The strain effects indicate that within the musculus strains the difference is less compared to 
This is an indication of evolution of the sexual dimorphism for the pattern of single foci positions.

The long bivalent pattern -- shows a different effect, both female and males have positions close to the center. This is a lack of telomeric pattern efpected for males

For short bivalents, we predict that the normalized position will be more medial

The short bivalent landscape pattern is more medial compared to males in all strains. However some strains have larger sex differences, WSB

@@

@@

```{r Q1.nrmF1CO.pos_M1, echo=FALSE}

lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)

lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )

lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )

anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)
anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)

lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Mouse.Table_BivData_4MM, REML = TRUE)

lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Mouse.Table_BivData_4MM )

lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Mouse.Table_BivData_4MM )

anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)
anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)

drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )

exactRLRT(lmer.Q1.nrm1CO_M1)

```{r Q1.Nrm.F1pos_M2, echo=FALSE}

#calq the means for normalized F1
#M2

Q1.glm_nrmF1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Mouse.Table_BivData_4MM)
#summary(Q1.glm_nrmF1pos_M2)

summry.Q1.glm_nrmF1pos_M2 <- summary(Q1.glm_nrmF1pos_M2)
round(summry.Q1.glm_nrmF1pos_M2$coefficients[,c(1,4)],5)

write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", sep=",",row.names = TRUE)

```{r Q1.nrmCO.pos_M3, echo=FALSE, eval=FALSE}

Q1.glm_nrmF1pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Mouse.Table_BivData_4MM)
#summary(Q1.glm_nrmF1pos_M2)

summry.Q1.glm_nrmF1pos_M3 <- summary(Q1.glm_nrmF1pos_M3)
round(summry.Q1.glm_nrmF1pos_M3$coefficients[,c(1,4)],5)

summry.Q1.glm_nrmF1pos_M3 <- summary(Q1.glm_nrmF1pos_M3)
round(summry.Q1.glm_nrmF1pos_M3$coefficients[,c(1,4)],5)

write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
```

The model results aren't as clear as I'd like.
For the Mol, both the sex effect and MOLF strain effect are significant (this means both female and male rec landscapes are effected - most of the time towards the middle

GENERALLY - Sex is the biggest effect for the 1CO landscape (female middle male telomereic)
molf
Not sure I understand how MSM is supper significant for 
There must be some strange effect due to the Mol subsp - MOLF isn't even in M2 (
Male is consistently significant across the two models.

Male and MOLF are the most significant effects for M3

@@

```{r Q1.t.test_1CO_pos, echo=FALSE, eval=FALSE}

#t.test for high and low group  -- logistic regression still need to format the high and low group

#mouse av
t.test_mouse.av.1CO_pos <- t.test(

  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$Rec.group == 1)],

  mouse.avs_4MM$norm.first.foci.pos[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc") &(mouse.avs_4MM$Rec.group == 0)] 
  )

#Single Biv Level
single.biv.1CO_pos_t.test <- t.test(

  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],

   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
  )

>>>>

```{r MLH1.FEMALE.TABLE, echo=FALSE}

@@

```{r MLH1.MALE.TABLE, echo=FALSE}
male_mouse.av <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

@@

```{r MLH1.MIXED.MODEL.setup, echo=FALSE, warning=FALSE, eval=TRUE}
#DF.HetC.MixedModel.HQ

@@

```{r CurBiv.Mouse.av.TABLE, echo=FALSE}

#this block should be in the data set up
#the mouse averages will be used

#assign high and low
#divide musc males into groups
mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))

@@

```{r totSC.viz4outliers, echo=FALSE}

@@

<!--
From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.

@@

skel -- takes the skeletonized midline of the bin objects
-->
# Main Figure

@@

The strain means for females were pretty similar. For female Domesticus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Dom")  )$mean_co` for _M. m. domesticus^WSB^_, _M. m. domesticus^G^_, and _M. m. domesticus$^LEW^_ respectively. In the three Musculus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Musc")  )$mean_co` for _M. m. musculus^PWD^_, _M. m. musculus^SKIVE^_, and _M. m. musculus^KAZ^_ respectively. In the molossinus strains the means are `r (main.table  %>% filter(sex == "female") %>% filter(subsp == "Mol")  )$mean_co` for _M. m. molossinus^ MSM^_ and _M. m. molossinus^MOLF^_ respectively.

While in two of the subspecies, the means for males show a greater range across means, notable in the mmmolossinus strains `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Mol")  )$mean_co` for _M. m. molossinus^MSM^_ and  _M. m. molossinus^MOLF^_ respectively. In the musculus the strain means are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Musc")  )$mean_co` for _M. m. musculus^PWD^_, _M. m. musculus^SKIVE^_, _M. m. musculus^KAZ^_, _M. m. musculus^TOM^_, _M. m. musculus^AST^_, and _M. m. musculus^CZECH^_ respectively. In the male Dom strains the means for MLH1 count per cell are `r (main.table  %>% filter(sex == "male") %>% filter(subsp == "Dom")  )$mean_co`  for  _M. m. domesticus^WSB^_, _M. m. domesticus^G^_, _M. m. domesticus^LEW^_ and _M. m. domesticus^PERC^_ respectively.

<!-- mouse level means-->

@@

# Analysis of Evolutionary Patterns

@@

Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
-->

@@

```{r female.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}

@@

#female strain table
#female_mouse.avs_for.analysis
F.strain.table <- ddply(female_mouse.avs_for.analysis, c("strain"), summarise,

                        Nmice = length(unique(mouse)),

                        total.cells  = sum(Ncells),

                        mouse.mean_co = mean(mean_co),

@@

                        mean.mouse_sd   = mean(sd, na.rm = TRUE ),

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se)
    )
F.strain.table

# I don't know why most of the table is blank
```

@@

In the dom  subspecies the ranges of mouse means are 
`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to

`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in _M. m. domesticus$^WSB^_ ,

`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` to

`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` in _M. m. domesticus$^G^_

@@

```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
#try customize the plot size

@@

                        #Standard error (mice vs cells)
                        mean.mouse_se  = mean(se)
    )
M.strain.table
#same issue as the female table .. ugh
```

@@

<!--
(discription of the nice boxplot figure)

@@

- the ratios of DMC1 : MLH1 counts!!
-->
```{r plot.stage.ttest, echo=FALSE, warning=FALSE}

@@

```{r DMC1_MLH1.cor, echo=FALSE, warning = FALSE}
#make category averages from AP_mouse_table.HQ
mean.MLH1.obs <- ddply(.data=AP_mouse_table.HQ,
                 .(category),
                 summarize, 
                 mean.MLH1 = mean(mean_co),
                 var.MLH1 = var(mean_co)
)

@@

PWD_DMC1.Z_CO_ratio <-( DMC1table.main %>% filter(category == "PWD male") )$mean.count.L / (DMC1table.main %>% filter(category == "PWD male"))$mean.MLH1

as.data.frame(DMC1table.main)

The table above indicates, there is not much variation across the MLH1 : DMC1 ratio (a proxy for the proportion of NCO).

The correlation with MLH1 and leptotene cells is `r MLH1.DMC1.L_cor`. The correlation with MLH1 and zygotene cells is `r MLH1.DMC1.Z_cor`.

`r c(WSB_DMC1.Z_CO_ratio, G_DMC1.Z_CO_ratio, MSM_DMC1.Z_CO_ratio, PWD_DMC1.Z_CO_ratio )` for  _M. m. domesticus^WSB^_ , _M. m. domesticus^G^_, _M. m. molossinus^MSM^_, and  _M. m. musculus^PWD^_ respecitively.

@@

@@

<!--
Since there is evidence for non-equal variance across strains for zygotene cells, so don't rely on lm()s that estimate the effect of strains.

The p-values from the t.tests of the high vs low groups, indicate that the high recombining are significantly higher for the L cells (p value = `r ttest.HighLow.L$p.value`) while the zygotene cells, there is not a significant difference across the high and low groups (value = `r ttest.HighLow.Z$p.value`).
-->

@@

ToDo:
- within strain across sex,
- males, across high and low groups
- look at mouse level proportions
-->

@@

@@

Q1.glm_totSC_M2 <- glm(mean.skel ~ subsp*strain * sex, data=Q1.lmer_totSC.mouse.av.table)
#summary(Q1.glm_totSC_M2)
#subsp effect, many strain, #male

@@

Q1.glm_totSC_M3 <- glm(mean.skel ~ strain * sex, data=Q1.lmer_totSC.mouse.av.table)
#summary(Q1.glm_totSC_M3)
#LOT OF SIGNIFICant strain effects in this model

@@

```{r Q1.SC.Length_M2, echo=FALSE}
Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)

@@

@@

@@

<!-- motivating

To quantify one of the major -- patterns that is different between male and female recombiantion landscapes
In order to test for evolution -- and examine sex differences in the recombination landscape (crossover position bias)
-->

we focus on the foci positions from one crossover (1CO) bivalents since the landscape patterns for multi crossover bivalents will be highly influences by crossover interference. We use the normalized foci positions (smaller meaning closer to the centromere and larger being cloerse to the telomere ). (all chromosomes in house mouse are acrocentric)

Below are the general patterns / plots -- short and long bivalent groups are shown to try and account for the chromsomes size effect.

@@

```{r Q1show.F1pos_Full, echo=FALSE}

Q1.mouse.av.biv_scatr_box   <- ggplot(mouse.avs_4MM, aes(x=sex, y=norm.first.foci.pos, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= norm.first.foci.pos ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \n all bivalents")
Q1.mouse.av.biv_scatr_box

```{r Q1_1COpos.t.test, echo=FALSE}
#testing,
#norm.first.foci.pos  - mean nrm.F1.pos for just 1CO bivalents

#WSB
Q1_1CO.pos_t.test.WSB <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "WSB") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "WSB") )$norm.first.foci.pos)
#G
Q1_1CO.pos_t.test.G <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "G") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "G") )$norm.first.foci.pos)
#LEW
Q1_1CO.pos_t.test.LEW <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "LEW") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "LEW") )$norm.first.foci.pos)

#PWD
Q1_1CO.pos_t.test.PWD <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "PWD") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "PWD") )$norm.first.foci.pos)

#KAZ
Q1_1CO.pos_t.test.KAZ <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "KAZ") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "KAZ") )$norm.first.foci.pos)

#SKIVE
Q1_1CO.pos_t.test.SKIVE <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "SKIVE") )$norm.first.foci.pos,
           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "SKIVE") )$norm.first.foci.pos )

#MSM
Q1_1CO.pos_t.test.MSM <- t.test((mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain == "MSM") )$norm.first.foci.pos,
                           (mouse.avs_4MM %>% filter(sex == "female") %>% filter(strain == "MSM") )$norm.first.foci.pos)

#Single Biv Level
#single.biv.1CO_pos_t.test <- t.test(

 # Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],

#   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
#  )

#t.tests for mous.av_nrm.1CO.pos 
```

In the domesticus strains the mouse averages for normalized single foci positions were significantly sexually dimorphic (with female)
are highly significantly more medially placed (t.test; p = `r Q1_1CO.pos_t.test.WSB$p.value`, `r Q1_1CO.pos_t.test.G$p.value`, and `r Q1_1CO.pos_t.test.LEW$p.value` for WSB, G and LEW respectively). With the musclus and molossinus strains displaying a similar pattern (t.test p = `r Q1_1CO.pos_t.test.PWD$p.value` and `r Q1_1CO.pos_t.test.MSM$p.value`) PWD and MSM. Two musculus  strains had a weaker signal, but it was still in the same direction (t.test; p =  `r Q1_1CO.pos_t.test.KAZ$p.value`, and `r Q1_1CO.pos_t.test.SKIVE$p.value` for KAZ and SKIVE respectively).

<!-- lmer and glm here -->

```{r Q1.CO1pos_lmer.M1, echo=FALSE}

#make sure this dataset has all the right strains
Q1.Foci.pos_DF <- Mouse.Table_BivData_4MM %>% filter(strain != "CZECH") %>% filter(strain != "MOLF")

lmer.Q1.nrm1CO_M1 <- lmer(norm.first.foci.pos ~ subsp * sex + (1|strain), data=Q1.Foci.pos_DF, REML = TRUE)

lmer.Q1.nrm1CO_M1reduce.sex <- lmer( norm.first.foci.pos ~ subsp  +(1|strain), data=Q1.Foci.pos_DF )

lmer.Q1.nrm1CO_M1reduce.subp <- lmer( norm.first.foci.pos ~ sex  +(1|strain), data=Q1.Foci.pos_DF )

lmer.Q1.nrm1CO_M1.sex <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.sex)$`Pr(>Chisq)`[2]
lmer.Q1.nrm1CO_M1.subp <-  anova(lmer.Q1.nrm1CO_M1, lmer.Q1.nrm1CO_M1reduce.subp)$`Pr(>Chisq)`[2]

lmer.Q1.nrm1CO_M1_interaction <- drop1(lmer.Q1.nrm1CO_M1, test="Chisq" )$`Pr(Chi)`[2]
lmer.Q1.nrm1CO_M1_strain <- exactRLRT(lmer.Q1.nrm1CO_M1)$p.value

lmer.Q1.nrm1CO_M1_results <- data.frame(sub.sp = lmer.Q1.nrm1CO_M1.subp, 
sex.results = lmer.Q1.nrm1CO_M1.sex , 
               interaction.pval = lmer.Q1.nrm1CO_M1_interaction,  rand.pval = lmer.Q1.nrm1CO_M1_strain)

lmer.Q1.nrm1CO_M1_results

```
For the Mixed model, sex and strain are significant effects for the normalized single foci position.

```{r Q1.CO1pos_glm.M2, echo=FALSE}

#in case factors aren't ordered
#Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
#Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
#Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)

Q1.glm_F1pos_M2 <- glm(norm.first.foci.pos ~ subsp * strain * sex, data=Q1.Foci.pos_DF)
summary(Q1.glm_F1pos_M2)

summry.Q1.glm_F1pos_M2 <- summary(Q1.glm_F1pos_M2)
round(summry.Q1.glm_F1pos_M2$coefficients[,c(1,4)],5)

#write.table(round(summry.Q1.glm_nrmF1pos_M2$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M2.csv", #sep=",",row.names = TRUE)

```{r Q1.CO1pos_glm.M3, echo=FALSE, eval=FALSE}

summry.Q1.glm_F1pos_M2

Q1.glm_1CO.pos_M3 <- glm(norm.first.foci.pos ~ strain * sex, data=Q1.Foci.pos_DF)
summary(Q1.glm_1CO.pos_M3)

summry.Q1.glm_1CO.pos_M3 <- summary(Q1.glm_1CO.pos_M3)
round(summry.Q1.glm_1CO.pos_M3$coefficients[,c(1,4)],5)

#write.table(round(summry.Q1.glm_nrmF1pos_M3$coefficients,5), "~./MLH1repo/summry.Q1.glm_nrmF1pos_M3.csv", sep=",",row.names = TRUE)
```

Lew and KAZ had slightly significant strain speicific effects across the two glms (these results will be explored more fully in the Q2 section). The general pattern that emerges across these models is the sex effect being the most significant, (with the male single foci position being more telomeric compared to females).

```{r Q1.1CO.pos_Sardell.eq, echo=FALSE, include=FALSE}
#IGNORE THIS FOR KNOW, THIS

#filter all Curated biv data for 1 strain
#WSB.Bivalents.DF , use single bivalent level observations, with the strain specific DFs

# the stardell paper sums the number of COs for bins (10)

#in my data it would be like adding -- 10 more colums, and marking 1 if there is a bin within it
#but then the model would be close to logistic

#summing the number of COs per bin, losses the chrom specific data....
#!! need to use a melted Data frame to capture all nrm.foci obs
#wsb.ttest<- mouse.avs_4MM %>% filter()
#start with same chrm classes
#the telomere dist and centromere dis were celulcate for the nearest foci ... but really these shhould be calq for each foci

WSB.Biv.1CO <- WSB.Bivalents.DF %>% filter(hand.foci.count == 1)

Sardell.mo.WSB <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex+ (1 | chromosomeLength), data = WSB.Biv.1CO)

PWD.Biv.1CO <- PWD.Bivalents.DF %>% filter(hand.foci.count == 1)

Sardell.mo.PWD <- lmer(PER_Foci_1 ~ dis.cent + telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
Sardell.mo.PWD_R.dis.cent <- lmer(PER_Foci_1 ~  telo_dist + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)
Sardell.mo.PWD_R.dis.telo <- lmer(PER_Foci_1 ~  dis.cent + sex + (1 | chromosomeLength), data = PWD.Biv.1CO)

anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.cent) #p= highly sig

anova(Sardell.mo.PWD,Sardell.mo.PWD_R.dis.telo) #p= v low

anova(Sardell.mo.PWD_R.dis.cent, Sardell.mo.PWD_R.dis.telo) #p = 1

#or AIC(model1,model2)

@@

@@
